preciseTAD: A machine learning framework for precise 3D domain boundary prediction at base-level resolution

### 2020-09-09 Manuscript/manuscript.Rmd
Add [@schreiber2019pitfall] ref

Random forest (RF) classification models (the `caret` v.6.0 R package [@kuhn2012caret]) were built to compare model performances between combinations of data resolutions, feature engineering procedures, and resampling techniques. A *holdout chromosome* technique was used for model evaluation on each autosomal chromosome. The $i^{th}$ holdout chromosome was identified and a data matrix, $A_{N \times (p+1)}$, was constructed by combining the binned genome from the remaining chromosomes ($1,2,\cdots, i-1,i+1,\cdots,21,22$), where $N=[n_{1} n_{2} \cdots n_{21} n_{22}]'$ and $n_{k}$ is the length of chromosome *k* after being binned into non-overlapping regions of resolution *r*, such that $k \ne i$. $p$, the number of annotations, together with the response vector **Y**, defined the column-wise dimension of the matrix $A$. Re-sampling was then performed on $A$, and an RF classifier was trained using 3-fold cross-validation to tune for the number of annotations to consider at each node (*mtry*). The number of trees (*ntree*) that were aggregated for each RF model was set to 500. The minimum number of observations per root node (*nodesize*) was set to .01% of the rows in the data. The binned data for the holdout chromosome $i$ was reserved for testing. Models were evaluated using Balanced Accuracy (BA), defined as the average of sensitivity and specificity:

@@

When evaluating performance across cell lines, a random forest model was built using Arrowhead/Peakachu-detected boundaries ("ground truth") and genomic annotations from one cell line such that the data matrix was constructed using genomic bins from $n-1$ chromosomes. The test data was constructed from the binned data of the holdout chromosome with genomic annotations from the other cell line. The predicted boundaries were evaluated for overlap with the "ground truth" boundaries from the that other cell line. This process was repeated for each holdout chromosome. To evaluate performance, we constructed receiver operating characteristic (ROC) curves composed of the average sensitivities and specificities and reported the area under the curve (AUC).

>>>>

Random forest (RF) classification models (the `caret` v.6.0 R package [@kuhn2012caret]) were built to compare model performances between combinations of data resolutions, feature engineering procedures, and resampling techniques. Following recommendations to evaluate the model on unseen data [@schreiber2019pitfall], a *holdout chromosome* technique was used for estimating model's performance. The $i^{th}$ holdout chromosome was identified and a data matrix, $A_{N \times (p+1)}$, was constructed by combining the binned genome from the remaining chromosomes ($1,2,\cdots, i-1,i+1,\cdots,21,22$), where $N=[n_{1} n_{2} \cdots n_{21} n_{22}]'$ and $n_{k}$ is the length of chromosome *k* after being binned into non-overlapping regions of resolution *r*, such that $k \ne i$. $p$, the number of annotations, together with the response vector **Y**, defined the column-wise dimension of the matrix $A$. Re-sampling was then performed on $A$, and an RF classifier was trained using 3-fold cross-validation to tune for the number of annotations to consider at each node (*mtry*). The number of trees (*ntree*) that were aggregated for each RF model was set to 500. The minimum number of observations per root node (*nodesize*) was set to .01% of the rows in the data. The binned data for the holdout chromosome $i$ was reserved for testing. Models were evaluated using Balanced Accuracy (BA), defined as the average of sensitivity and specificity:

@@

When evaluating performance across cell lines, we followed the same strategy of evaluating a model trained in one cell line on unseen data from another cell line [@schreiber2019pitfall]. A random forest model was built using Arrowhead/Peakachu-detected boundaries ("ground truth") and genomic annotations from one cell line such that the data matrix was constructed using genomic bins from $n-1$ chromosomes. The test data was constructed from the binned data of the holdout chromosome with genomic annotations from the other cell line. The predicted boundaries were evaluated for overlap with the "ground truth" boundaries from the that other cell line. This process was repeated for each holdout chromosome. To evaluate performance, we constructed receiver operating characteristic (ROC) curves composed of the average sensitivities and specificities and reported the area under the curve (AUC).
